\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage{parskip}
\usepackage{tabularx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{geometry}
\usepackage{booktabs}
\usepackage{centernot}
\usepackage{hyperref}
\usepackage{eufrak}
\usepackage{graphicx}
\graphicspath{{pics/}}
\geometry{a4paper, left=20mm, right=20mm, top=20mm, bottom=20mm}

\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{Anthony Catterwell}
\chead{\textsc{University of Edinburgh}}
\rhead{Fundamentals of Optimization}

\title{Fundamentals of Optimization Notes}
\author{Anthony Catterwell}

\pagenumbering{gobble}

\begin{document}
\maketitle
\tableofcontents

\section{Lecture 1}

\subsection{LP Formulation}
\subsection{Graphs \& Directed Graphs}
\subsection{Network Flow Problem}
\subsection{Quadratic Programming}


\section{Lecture 2}

\subsection{Convexity}
\begin{itemize}
    \item \textbf{Definition} \emph{Convexity} \\
        A set $C \subset \mathbb{R}^n$ is convex
        $\iff \forall x, y \in C, \forall \lambda \in [0, 1] \ f(\lambda x + (1 - \lambda))
        \leq \lambda f(x) + (1 - \lambda) f(y)$

    \item \textbf{Definition} \emph{Epigraph} \\
        The epigraph of a function is the set lying \textbf{above} the function.\
        i.e.\ If $C$ is a subset of $\mathbb{R}^n$, and function $f : C \mapsto \mathbb{R}$,
        then the epigraph of $f$ is
        \[
            \mathrm{epi}(f) = \{ (x, z) | x \in C, z \geq f(x) \} \in \mathbb{R}^{n+1}
        \]
\end{itemize}

\begin{itemize}
    \item \textbf{Definition} \emph{Convex set} \\
        A set $C \subset \mathbb{R}^n$ is convex
        \[
            \iff \forall x, y \in C, \forall \lambda \in [0, 1], \quad
            \exists z = \lambda x + (1 - \lambda) y \in C
        \]

    \item \textbf{Definition} Function $f : C \mapsto \mathbb{R}$ is \emph{convex} if and only if
        \[
            f(\lambda x + (1 - \lambda) y) \leq \lambda f(x) + (1-\lambda) f(y), \
            \forall x, y \in C, \forall \lambda \in [0, 1]
        \]

    \item \textbf{Definition} Function $f : C \mapsto \mathbb{R}$ is \emph{concave} if and only if
        \[
            f(\lambda x + (1 - \lambda) y) \geq \lambda f(x) + (1-\lambda) f(y), \
            \forall x, y \in C, \forall \lambda \in [0, 1]
        \]

    \item \textbf{Definition} Function $f : C \mapsto \mathbb{R}$ is \emph{strictly convex} if and only if
        \[
            f(\lambda x + (1 - \lambda) y) < \lambda f(x) + (1-\lambda) f(y), \
            \forall x, y \in C, \forall \lambda \in [0, 1]
        \]

    \item \textbf{Definition} Function $f : C \mapsto \mathbb{R}$ is \emph{strictly concave} if and only if
        \[
            f(\lambda x + (1 - \lambda) y) > \lambda f(x) + (1-\lambda) f(y), \
            \forall x, y \in C, \forall \lambda \in [0, 1]
        \]

    \item \textbf{Definition} A vector $\vec{x}$ is a \textbf{local} minimum of $f$ if
        \[
            \exists \epsilon > 0 \ \text{such that} \
            f(\vec{x}) \leq f(x), \forall x |x - \vec{x}| < \epsilon
        \]

    \item \textbf{Definition} A vector $\vec{x}$ is a \textbf{global} minimum of $f$ if
        \[
            f(\vec{x}) \leq f(x), \ \forall x \in X
        \]

    \item If $X$ is a convex set and $f : X \mapsto \mathbb{R}$ is a convex function,
        then the minimization problem is said to be \emph{convex}.

    \item \textbf{Lemma} In a convex minimization problem a \textbf{local} minimum is a
        \mathbf{global} minimum.

    \item \textbf{Definition} A linear map $L : \mathbb{R}^n \to \mathbb{R}^m$ satisfies
        \begin{enumerate}
            \item $L(x + y) = L(x) + L(y)$
            \item $\alpha L(x) = L(\alpha x)$ for scalar $\alpha$.
        \end{enumerate}

    \item \textbf{Definition} An affine map is a linear map \emph{plus} a shift of the origin.
        It takes the form $F(x) = d + Mx$, where $d \in \mathbb{R}^m$.

    \item \textbf{Lemma} An affine function (and hence also a linear function) is both
        convex and concave.

\end{itemize}

\subsection{Properties of Convex Sets and Functions}
\begin{enumerate}
    \item For any collection $\{ C_i | i \in I \}$ of convex sets, the intersection
        $\cap_{i \in I} C_i$ is convex.

    \item The vector sum $\{ x_1 + x_2 | x_1 \in C_1, x_2 \in C_2 \}$
        of two convex sets $C_1$ and $C_2$ is convex.

    \item The image of a convex set under a linear map is convex.\
        i.e. If $C$ is a convex set and $L$ a linear map $\{ L(x) | x \in C \}$ is convex.
        (True also for affine map.)

    \item For $C$ a convex set,
        $f : C \mapsto \mathbb{R}$ a convex function and $L : C \mapsto \mathbb{R}$
        and affine function, the \textbf{level sets}
        $\{ x \in C | f(x) \leq \alpha \}$
        and
        $\{ x \in C | L(x) = \alpha \}$
        are convex for all scalars $\alpha$.

    \item For $C$ a convex set and any collection $f_i : C \mapsto \mathbb{R} | i \in I$ of convex
        functions, a non-negative weighted sum, i.e.
        $f = \sum_{i \in I} w_f f_i : C \mapsto \mathbb{R}$,
        where $w_i \geq 0$, $i \in I$, is convex.

    \item For $I$ a (finite) index set, $C$ a convex set, and $f_i : C \mapsto \mathbb{R}$
        a convex function $\forall i \in I$, the function $h : C \mapsto \mathbb{R}$ defined by
        $h(x) = \max_{i \in I} f_i(x)$ is convex.

    \item For $C$ a convex set, $f: C \mapsto \mathbb{R}$ a convex function, and
        $g : \mathbb{R} \mapsto \mathbb{R}$
        a convex and non-decreasing function, the function $h : C \mapsto \mathbb{R}$ defined
        by $h(x) = g(f(x))$ is convex.

    \item Let $C \subseteq \mathbb{R}^n$ be a convex set and $f : C \mapsto \mathbb{R}$ be
        differentiable over $C$.
        \begin{enumerate}
            \item The function $f$ is convex if and only if
                \[
                    f(y) \geq f(x) + \nabla^T f(x)(y-x), \quad \forall x, y \in C
                \]

            \item if the inequality is strict $\forall x \neq y$, then $f$ is strictly convex.
        \end{enumerate}

    \item Let $C \subseteq \mathbb{R}^n$ be a convex set and $f : C \mapsto \mathbb{R}$
        be twice continuously differentiable over $C$.
        \begin{enumerate}
            \item If $f$ is convex in $C$ and $C$ has an interior (i.e.\ a point not on its boundary),
                then $\nabla^2 f(x)$ is positive semi-definite at all $x \in C$.

            \item If $\nabla^2 f(x)$ is positive semi-definite for all $x \in C$, then $f$ is convex in $C$.

            \item If $\nabla^2 f(x)$ is positive definite for all $x \in C$, then $f$ is strictly convex.
        \end{enumerate}

    \item Let $H$ be a square symmetric matrix and let $f(x) = x^T H x$
        (so $f$ is a quadratic function).
        \begin{enumerate}
            \item $f$ is convex if and only if $H$ is positive semi-definite.

            \item $f$ is strictly convex if and only if $H$ is positive definite.
        \end{enumerate}
\end{enumerate}

\end{document}
